---
title: "PAMGuide calibraition"
format: html
editor: visual
---

```{r, warning=FALSE}
source('PAMGuide.R')

library(tidyverse)
library(readxl)
library(tuneR)
library(seewave)
library(gsignal)
library(pracma)

write_results <- TRUE

#Set up folder structures
audio_data_folder <- "G:/Drive condivisi/Sound_recorder_calibration_DB/data/Audio/"
slm_data_folder <- "G:/Drive condivisi/Sound_recorder_calibration_DB/data/SLM/raw/"
output_folder <- "G:/Drive condivisi/Sound_recorder_calibration_DB/output"
output_file <- "calibration_database"
```

Steps for calibration:

1.  Extract reference value from SLM and transform to linear

2.  Estimate 1 kHz levels from AudioMoth recordings

    a\. Use bespoke TOL function to estimate dB levels at 1 kHz

    b\. Calculate RMS and tranform to linear

3.  Compute sensitivity factor ready to go in PAMGuide

# 1. Create final database

```{r}
df <- data.frame(sensor = NA,
                 calibration_test = NA,
                 sensor_dB = NA,
                 slm_dB = NA,
                 calibration_offset = NA)
```

# 2. Calculate per sensor calibration value

Bespoke TOL function:

```{r}
third_octave_levels <- function(audio, fs, nfft=4096){
  
  #Step 1: Welch PSD
  spec <- pwelch(
    audio, 
    window = hanning(nfft),
    overlap = 0.5,
    nfft = nfft,
    fs = fs
  )
  
  freqs <- spec$freq
  psd <- spec$spec
  
  #Step 2: third-octave center frequency
  center_freqs <- c(25,31.5,10,50,63,80, 100,125,160,200,250,315,400,500,630,800,1000,1250,1600,200,2500,3150,4000,5000,6300,8000,10000,12500,16000,20000)
  
  #Step 3: Band edges
  k <- 2^(1/6)
  lower <- center_freqs/k
  upper <- center_freqs*k
  
  #Step 4: Integrate PSD in each band
  levels <- NA
  for(i in seq_along(center_freqs)){
    idx <- which(freqs >= lower[i] & freqs <= upper[i])
    
    if(length(idx) >0){
      #band_power <- pracma::trapz(freqs[idx], psd[idx])
      bandwidth = max(freqs[idx]) - min(freqs[idx]) 
      band_power <- sum(psd[idx]) * bandwidth
      levels[i] <- 10 + log10(band_power)
    } else {
      levels[i] <- NA
    }
  }
  
  return(list(freq = center_freqs, levels_df = levels))
}

average_dB <- function(levels_dB) {
  linear <- 10^(levels_dB / 10)
  mean_linear <- mean(linear, na.rm = TRUE)
  mean_dB <- 10 * log10(mean_linear)
  return(mean_dB)
}

```

Calibration test metadata:

```{r}

calib_test <- "Calib0035"
sensor <- "AMO84"

use_all_three_tests = TRUE

date_test <- "2026-03-27"

# SLM reference test
test_time <- "10:43:38"
slm_test <- "043"
# slm_time_mismatch <- 16 #seconds Introduced code to estimate this from the "clap" sound at the beginning of the test, still need to check if it works for all the tests!

WildSOUND_calibration <- FALSE #TRUE if the short audio file was used
```

Next bit of the code:

1.  Calculate TOL
2.  Extract dB values from the SLM
3.  Calculate calibration offset
4.  Write out results

```{r}

df_row <- df[1,]

########################################
#  1. Run TOL analysis and extract mean dB level recorded
#######################################

if(use_all_three_tests == TRUE){
  kHz1_tests <- c("_1kHz_a","_1kHz_b","_1kHz_c")
} else {
  print("Manually specify tests to be used")
  break
}
  
for(x in 1:length(kHz1_tests)){
  #Read wav
  wav <- readWave(paste0(audio_data_folder, calib_test, "/", calib_test,"_", sensor, kHz1_tests[x],".wav"))
    
  #Convert to numeric vector
  audio <- wav@left
  fs <- wav@samp.rate
    
  #Run analysis 
  result <- third_octave_levels(audio, fs)
    
  #View results
  dat.manual <- data.frame(
    Frequency = result$freq,
    Level_dB = result$levels_df
  )
  
  if(x == 1){
    sensor_dB <- dat.manual$Level_dB[which(dat.manual$Frequency == 1000)]
  } else {
    sensor_dB <- c(sensor_dB, dat.manual$Level_dB[which(dat.manual$Frequency == 1000)])
  }

}

#Calculate mean value across the three tests
SPL_linear <- 10^(sensor_dB/10)
mean_SPL_linear <- mean(SPL_linear)
mean_SPL_dB <- 10*log10(mean_SPL_linear)

#Append results to database
df_row$sensor <- sensor
df_row$calibration_test <- calib_test
df_row$sensor_dB <- mean_SPL_dB

########################################
#  2. Extract dB values from the SLM
#######################################

slm_file <- list.files(paste0(slm_data_folder, gsub("-","",date_test)), pattern=paste0(slm_test,".xlsx"), full.names = TRUE)

slm_dat <- readxl::read_excel(slm_file, sheet = "Profilo storico") %>% select("Tempo", "1/3 LZFeq 1000") %>% rename(SPL_1kHz = "1/3 LZFeq 1000")

# Find the clap of the beginning of the test to estimate any potential time mismatch. Use the difference in SPL to identify sudden peaks and then assume that is the closest one to the recorded test time
test_time <- paste0(date_test," ",test_time) %>% as.POSIXct(., tz="UTC")
  
slm_dat <- slm_dat %>% mutate(diff = SPL_1kHz - lag(SPL_1kHz),
                              time_to_test = difftime(Tempo, test_time) %>% as.numeric())
  
clap <- slm_dat %>%
  arrange(desc(diff)) %>%
  slice(1:30)
n <- which.min(abs(clap$time_to_test))
  
time <- clap$Tempo[n]

if(WildSOUND_calibration == TRUE){
  time1_start <- time + 13
  time1_end <- time + 32
    
  time2_start <- time + 46 
  time2_end <- time + 65
    
  time3_start <- time + 79
  time3_end <- time + 98
} else {
    time1_start <- time + 108
    time1_end <- time + 127
    
    time2_start <- time + 140
    time2_end <- time + 159
    
    time3_start <- time + 172
    time3_end <- time + 191

}
  
slm1 <- slm_dat %>% dplyr::filter(Tempo >= time1_start & Tempo <= time1_end) 
slm2 <- slm_dat %>% dplyr::filter(Tempo >= time2_start & Tempo <= time2_end)
slm3 <- slm_dat %>% dplyr::filter(Tempo >= time3_start & Tempo <= time3_end)

x_slm <- c(average_dB(slm1$SPL_1kHz), average_dB(slm2$SPL_1kHz), average_dB(slm3$SPL_1kHz))
x_slm <- average_dB(x_slm)

df_row$slm_dB <- x_slm

########################################
#  3. Calculate calibration offset
#######################################

offset <- x_slm - mean_SPL_dB
df_row$calibration_offset <- offset

########################################
#  4. Write out results
#######################################

df <- rbind(df, df_row)
df <- df %>% dplyr::filter(!is.na(sensor))

#if(file.exists(paste0(output_folder,"/",output_file,".csv"))){
#  df1 <- read.csv(paste0(output_folder,"/",output_file,".csv"))
#  df1 <- rbind(df1, df)
#  write.csv(df1, paste0(output_folder,"/",output_file,".csv")) 
#} else {
# write.csv(df, paste0(output_folder,"/",output_file,".csv")) 
#}

```

Next things to check:

-   Problem with AMO92 test extraction that needs to be fixed

-   Try to run PAMGuide with and without sensitivity correction to compare the results

-   Start processing the offset of all the devices tested.

Potential improvements to the code:

-   Add read in the calibration database and extract the relevant info for all the files that have extracted audio

-   Cross-reference with calibration output to check which ones have been processed already

-   Update the offset extraction in the calibration database
