---
title: "PAMGuide calibraition"
format: html
editor: visual
author: Virginia Iorio
---

```{r, warning=FALSE}
library(tidyverse)
library(readxl)
library(tuneR)
library(seewave)
library(gsignal)
library(pracma)

write_results <- TRUE

#Set up folder structures
audio_data_folder <- "G:/Drive condivisi/Sound_recorder_calibration_DB/data/Audio/"
slm_data_folder <- "G:/Drive condivisi/Sound_recorder_calibration_DB/data/SLM/raw/"
output_folder <- "G:/Drive condivisi/Sound_recorder_calibration_DB/output"
output_file <- "calibration_offset_database"
```

Steps for calibration:

1.  Extract reference value from SLM and transform to linear

2.  Estimate 1 kHz levels from AudioMoth recordings

    a\. Use bespoke TOL function to estimate dB levels at 1 kHz

    b\. Calculate RMS and tranform to linear

3.  Compute sensitivity factor ready to go in PAMGuide

# Calculate per sensor calibration value

Bespoke TOL function:

```{r}
third_octave_levels <- function(audio, fs, nfft=4096){
  
  #Step 1: Welch PSD
  spec <- pwelch(
    audio, 
    window = hanning(nfft),
    overlap = 0.5,
    nfft = nfft,
    fs = fs
  )
  
  freqs <- spec$freq
  psd <- spec$spec
  
  #Step 2: third-octave center frequency
  center_freqs <- c(25,31.5,10,50,63,80, 100,125,160,200,250,315,400,500,630,800,1000,1250,1600,200,2500,3150,4000,5000,6300,8000,10000,12500,16000,20000)
  
  #Step 3: Band edges
  k <- 2^(1/6)
  lower <- center_freqs/k
  upper <- center_freqs*k
  
  #Step 4: Integrate PSD in each band
  levels <- NA
  for(i in seq_along(center_freqs)){
    idx <- which(freqs >= lower[i] & freqs <= upper[i])
    
    if(length(idx) >0){
      #band_power <- pracma::trapz(freqs[idx], psd[idx])
      bandwidth = max(freqs[idx]) - min(freqs[idx]) 
      band_power <- sum(psd[idx]) * bandwidth
      levels[i] <- 10 + log10(band_power)
    } else {
      levels[i] <- NA
    }
  }
  
  return(list(freq = center_freqs, levels_df = levels))
}

average_dB <- function(levels_dB) {
  linear <- 10^(levels_dB / 10)
  mean_linear <- mean(linear, na.rm = TRUE)
  mean_dB <- 10 * log10(mean_linear)
  return(mean_dB)
}

```

Find SLM reference test rows:

-   due to the SLM clock not being always synchronized need to manually check for the clap to correct for the time lag.

-   Load the file

-   Manually check for the clap, it should be a dB peak above the 30 dB, or look for roughly 100 (test with Julia) or 10 (WildSOUND calibration) lines before the actual tests start (i.e. consistent high levels of dB).

```{r}
date_test <- "2026-03-19"
slm_test <- "038"

slm_file <- list.files(paste0(slm_data_folder, gsub("-","",date_test)), pattern=paste0(slm_test,".xlsx"), full.names = TRUE)

slm_dat <- readxl::read_excel(slm_file, sheet = "Profilo storico") %>% select("Tempo", "1/3 LZFeq 1000") %>% rename(SPL_1kHz = "1/3 LZFeq 1000")
```

Calibration test metadata:

```{r}
calib_test <- "Calib0029"
sensor <- "SM007"

use_all_three_tests = TRUE

date_test <- "2026-03-19"

# SLM reference test
test_time <- "10:25:47"
slm_test <- "038"

WildSOUND_calibration <- FALSE #TRUE if the short audio file was used
```

Next bit of the code:

1.  Calculate TOL
2.  Extract dB values from the SLM
3.  Calculate calibration offset
4.  Write out results

```{r}

df <- data.frame(sensor = NA,
                 calibration_test = NA,
                 sensor_dB = NA,
                 slm_dB = NA,
                 calibration_offset = NA)

########################################
#  1. Run TOL analysis and extract mean dB level recorded
#######################################

if(use_all_three_tests == TRUE){
  kHz1_tests <- c("_1kHz_a","_1kHz_b","_1kHz_c")
} else {
  print("Manually specify tests to be used")
  break
}
  
for(x in 1:length(kHz1_tests)){
  #Read wav
  wav <- readWave(paste0(audio_data_folder, calib_test, "/", calib_test,"_", sensor, kHz1_tests[x],".wav"))
    
  #Convert to numeric vector
  audio <- wav@left
  fs <- wav@samp.rate
    
  #Run analysis 
  result <- third_octave_levels(audio, fs)
    
  #View results
  dat.manual <- data.frame(
    Frequency = result$freq,
    Level_dB = result$levels_df
  )
  
  if(x == 1){
    sensor_dB <- dat.manual$Level_dB[which(dat.manual$Frequency == 1000)]
  } else {
    sensor_dB <- c(sensor_dB, dat.manual$Level_dB[which(dat.manual$Frequency == 1000)])
  }

}

#Calculate mean value across the three tests
SPL_linear <- 10^(sensor_dB/10)
mean_SPL_linear <- mean(SPL_linear)
mean_SPL_dB <- 10*log10(mean_SPL_linear)

#Append results to database
df$sensor <- sensor
df$calibration_test <- calib_test
df$sensor_dB <- mean_SPL_dB

########################################
#  2. Extract dB values from the SLM
#######################################

slm_file <- list.files(paste0(slm_data_folder, gsub("-","",date_test)), pattern=paste0(slm_test,".xlsx"), full.names = TRUE)

slm_dat <- readxl::read_excel(slm_file, sheet = "Profilo storico") %>% select("Tempo", "1/3 LZFeq 1000") %>% rename(SPL_1kHz = "1/3 LZFeq 1000")

test_time <- paste0(date_test," ",test_time) %>% as.POSIXct(., tz="UTC")

if(WildSOUND_calibration == TRUE){
  time1_start <- test_time + 13
  time1_end <- test_time + 32
    
  time2_start <- test_time + 46 
  time2_end <- test_time + 65
    
  time3_start <- test_time + 79
  time3_end <- test_time + 98
} else {
    time1_start <- test_time + 108
    time1_end <- test_time + 127
    
    time2_start <- test_time + 140
    time2_end <- test_time + 159
    
    time3_start <- test_time + 172
    time3_end <- test_time + 191

}
  
slm1 <- slm_dat %>% dplyr::filter(Tempo >= time1_start & Tempo <= time1_end) 
slm2 <- slm_dat %>% dplyr::filter(Tempo >= time2_start & Tempo <= time2_end)
slm3 <- slm_dat %>% dplyr::filter(Tempo >= time3_start & Tempo <= time3_end)

x_slm <- c(average_dB(slm1$SPL_1kHz), average_dB(slm2$SPL_1kHz), average_dB(slm3$SPL_1kHz))
x_slm <- average_dB(x_slm)

df$slm_dB <- x_slm

########################################
#  3. Calculate calibration offset
#######################################

offset <- x_slm - mean_SPL_dB
df$calibration_offset <- offset

########################################
#  4. Write out results
#######################################

if(file.exists(paste0(output_folder,"/",output_file,".csv"))){
  df1 <- read.csv(paste0(output_folder,"/",output_file,".csv"))
  df1 <- rbind(df1, df)
  write.csv(df1, paste0(output_folder,"/",output_file,".csv"), row.names = FALSE) 
} else {
 write.csv(df, paste0(output_folder,"/",output_file,".csv"), row.names = FALSE) 
}

```

Next things to check:

-   Problem with AMO92 test extraction that needs to be fixed

-   Try to run PAMGuide with and without sensitivity correction to compare the results

-   Start processing the offset of all the devices tested.

Potential improvements to the code:

-   Add read in the calibration database and extract the relevant info for all the files that have extracted audio

-   Cross-reference with calibration output to check which ones have been processed already

-   Update the offset extraction in the calibration database
