---
title: "PAMGuide calibraition"
format: html
editor: visual
author: Virginia Iorio
---

```{r, warning=FALSE, message = FALS}
library(tidyverse)
library(readxl)
library(tuneR)
library(seewave)
library(gsignal)
library(pracma)

write_results <- TRUE

#Set up folder structures
audio_data_folder <- "G:/Drive condivisi/Sound_recorder_calibration_DB/data/Audio/"
slm_data_folder <- "G:/Drive condivisi/Sound_recorder_calibration_DB/data/SLM/raw/"
output_folder <- "G:/Drive condivisi/Sound_recorder_calibration_DB/output"
output_file <- "calibration_offset_database2"

#Averag dB function used to avergae across dB transforming them first into linear values.
average_dB <- function(levels_dB) {
  linear <- 10^(levels_dB / 10)
  mean_linear <- mean(linear, na.rm = TRUE)
  mean_dB <- 10 * log10(mean_linear)
  return(mean_dB)
}
```

#### Overview of calibration steps

In the lab a type 1 Sound Level Meter (slm) and an acoustic sensor (either AudioMoth or SongMeter Micro 2) were exposed to three playbacks of a 1 kHz tonal sound. The trials were conducted at a fixed geometry to allow calibration by substitution.

The following code is used to:

1.  Find SLM reference row (manual step).

2.  Specify calibration metadata (manual step).

3.  Extract linear RMS amplitude from the raw waveform for each trial of the acoustic sensor, applying a 1 kHz band pass filter.

4.  RMS values are averaged in the linear domain.

5.  Extract slm average dB during each trial and across trials (dB are transformed on the linear scale prior to averaging).

6.  Calibration offset is derived by:

    *calibration_offset_dB \<- slm_dB - 20 \* log10(sensor_rms)*

# Calculate per sensor calibration value

## 1. Find SLM reference test rows

-   due to the SLM clock not being always synchronized need to manually check for the clap to correct for the time lag.

-   Load the file

-   Manually check for the clap, it should be a dB peak above the 30 dB, or look for roughly 100 (test with Julia) or 10 (WildSOUND calibration) lines before the actual tests start (i.e. consistent high levels of dB).

```{r}
date_test <- "2026-04-13"
slm_test <- "091"

slm_file <- list.files(paste0(slm_data_folder, gsub("-","",date_test)), pattern=paste0(slm_test,".xlsx"), full.names = TRUE)

slm_dat <- readxl::read_excel(slm_file, sheet = 5)

# Adjust for some outputs being in Italian and others in English
if(colnames(slm_dat)[4] == "Tempo"){
  slm_dat <- slm_dat %>% select("Tempo", "1/3 LZFeq 1000") %>% rename(SPL_1kHz = "1/3 LZFeq 1000")
} else {
  slm_dat <- slm_dat %>% select("Time", "1/3 LZFeq 1000") %>% rename(SPL_1kHz = "1/3 LZFeq 1000", Tempo = "Time")
}
```

## 2. Insert calibration test metadata

```{r}
calib_test <- "Calib0059"
sensor <- "SM021"

use_all_three_tests = TRUE

date_test <- "2026-04-09"

# SLM reference test
test_time <- "10:16:13"
slm_test <- "079"

WildSOUND_calibration <- TRUE #TRUE if the short audio file was used
```

## 3. Calculate calibration offset

```{r}

df <- data.frame(sensor = NA,
                 calibration_test = NA,
                 digital_rms = NA,
                 slm_dB = NA,
                 slm_pa = NA,
                 sensitivity_pa_per_unit = NA,
                 calibration_offset_dB = NA)

########################################
#  1. Run TOL analysis and extract mean dB level recorded
#######################################

if(use_all_three_tests == TRUE){
  kHz1_tests <- c("_1kHz_a","_1kHz_b","_1kHz_c")
} else {
  print("Manually specify tests to be used")
  break
}
  
for(x in 1:length(kHz1_tests)){
  #Read wav
  wav <- readWave(paste0(audio_data_folder, calib_test, "/", calib_test,"_", sensor, kHz1_tests[x],".wav"))
    
  #Extract linear RMS amplitude from the raw waveform for each trial.
  audio <- wav@left / (2^(wav@bit - 1))
  fs <- wav@samp.rate
  
  #Apply 1 kHz band pass filter
  bp <- butter(
    n = 4,
    w = c(700, 1300) / (fs / 2),
    type = "pass"
    )
  
  #Filter the data
  audio_filt <- filtfilt(bp, audio)
  
  digital_rms_test <- sqrt(mean(audio_filt^2))
  
  if(x == 1){
    digital_rms <- digital_rms_test
  } else {
    digital_rms <- c(digital_rms, digital_rms_test)
  }

}

#Calculate mean value across the three tests in the linear domain
digital_rms <- mean(digital_rms)

#Append results to database
df$sensor <- sensor
df$calibration_test <- calib_test
df$digital_rms <- digital_rms

########################################
#  2. Extract value from the SLM
#######################################

slm_file <- list.files(paste0(slm_data_folder, gsub("-","",date_test)), pattern=paste0(slm_test,".xlsx"), full.names = TRUE)

slm_dat <- readxl::read_excel(slm_file, sheet = 5)

# Adjust for some outputs being in Italian and others in English
if(colnames(slm_dat)[4] == "Tempo"){
  slm_dat <- slm_dat %>% select("Tempo", "1/3 LZFeq 1000") %>% rename(SPL_1kHz = "1/3 LZFeq 1000")
} else {
  slm_dat <- slm_dat %>% select("Time", "1/3 LZFeq 1000") %>% rename(SPL_1kHz = "1/3 LZFeq 1000", Tempo = "Time")
}

test_time <- paste0(date_test," ",test_time) %>% as.POSIXct(., tz="UTC")

if(WildSOUND_calibration == TRUE){
  time1_start <- test_time + 13
  time1_end <- test_time + 32
    
  time2_start <- test_time + 46 
  time2_end <- test_time + 65
    
  time3_start <- test_time + 79
  time3_end <- test_time + 98
} else {
    time1_start <- test_time + 108
    time1_end <- test_time + 127
    
    time2_start <- test_time + 140
    time2_end <- test_time + 159
    
    time3_start <- test_time + 172
    time3_end <- test_time + 191
}
  
# Extract the recorded values during the three tests
slm1 <- slm_dat %>% dplyr::filter(Tempo >= time1_start & Tempo <= time1_end) 
slm2 <- slm_dat %>% dplyr::filter(Tempo >= time2_start & Tempo <= time2_end)
slm3 <- slm_dat %>% dplyr::filter(Tempo >= time3_start & Tempo <= time3_end)

# Avergae dB across the three tests
slm_dB <- c(average_dB(slm1$SPL_1kHz), average_dB(slm2$SPL_1kHz), average_dB(slm3$SPL_1kHz))
slm_dB <- average_dB(slm_dB)
slm_pa <- 20e-6 * 10^(slm_dB / 20)

#Append to database slm value both in dB and Pa value
df$slm_dB <- slm_dB
df$slm_pa <- slm_pa

########################################
#  3. Calculate calibration offset
#######################################

df$sensitivity_pa_per_unit <- slm_pa/digital_rms
df$calibration_offset_dB <- slm_dB - 20 * log10(digital_rms)

########################################
#  4. Write out results
#######################################

df

if(file.exists(paste0(output_folder,"/",output_file,".csv"))){
  df1 <- read.csv(paste0(output_folder,"/",output_file,".csv"), sep=",")
  df1 <- rbind(df1, df)
  write.csv(df1, paste0(output_folder,"/",output_file,".csv"), row.names = FALSE) 
} else {
 write.csv(df, paste0(output_folder,"/",output_file,".csv"), row.names = FALSE) 
}

```
